## Trishal Varma
## Hw1 - Machine Learning


```{r}
data(Boston)
str(Boston)
```

#Step 1 
# The boston data sheet contains infomration about the city of Boston. 
#It contains things like Crime, businesses, nitrogen oxide concentration, 
#and information on tax valuation of properties. 

#Step 2 

```{r}
head(Boston, 5)
tail(Boston, 2)
Boston[5,]
head(Boston, 5)[1]
names(Boston)
```

#Step 3
```{r}
mean(Boston$crim)
median(Boston$crim)
range(Boston$crim)
```

#Step 4
```{r}
hist(Boston$crim,
     main = "Histogram of Boston Crime",
     xlab = "Crime",
     xlim=c(0,60),
     col = "red")
```

#The histogram shows the crime per capita in the city of Boston, and the 
#frequency at which it is commited. 

#Step 5
```{r}
cor.test(Boston$crim, Boston$medv)
```

#Step 6
```{r}
plot(x=Boston$rm,
     y=Boston$medv,
     main = "Boston Housing Based on Income",
     xlab = "Available rooms",
     ylab = "Median income",
     col.lab = "red",
    col = "darkslategrey")
cor.test(Boston$rm, Boston$medv)
```

#As the median value increases, the number of rooms available increases. 
#This is a positive correlation.

#Step 7
```{r}
plot(x=Boston$chas,
     y=Boston$medv,
     main = "River factor in value",
     xlab = "River bound",
     ylab = "Median income")
cor.test(Boston$chas, Boston$medv)
```

# The Charles river also shows a positive correlation, 
#meaning that the houses bound by the river have higher median house value, 
#hence the positive correlation. 

#Step 8
```{r}
summary(Boston$rad) # Summary provides the min, median, mean, and max of the 
```
#accessability to the raidial highways. 
```{r}
unique(Boston$rad) # Unique removes dupicates from the value. Showing the 
```
# signifacnt numbers that are available. 

```{r}
sum(Boston$rad == 24)
mean((132/506) *100) #26.09% of the homes make up rad access of 24 homes. 
```

#Step 9 

```{r}
far <- Boston$rad
ifelse(far == 24,"TRUE","FALSE")

plot(x=far,
     y=Boston$medv,
     main = "Highway based on Median Value of Homes",
     xlab = "Radian Highway",
     ylab = "Median Value of Homes",
     col = "darkslategrey",
     col.lab = "dodgerblue2")
```

# The graph shows the highway access of homes that have access to the highway. 

#Step 10 
```{r}
cols <- c("crim","rm","lstat","medv")
summary(Boston[cols])
which.max(Boston$medv)
Boston[which.max(Boston$medv), cols]
```
