---
Name: Trishal Varma
Class: 4375 - Intro to Machine Learning
Homework: 5
output: pdf_document
---

#### Problem 1 ####

# Step 1 
Setting up the data Default using the ISLR library. Then running dim and names to get more information.
Seed function to replicate the same data, and then dividing into test and train.
```{r}
#a
library(ISLR)
data(Default)
#b
dim(Default)
names(Default)
#c
df <- Default
set.seed(1234)
#d
smp_size <- sample(1:nrow(df), .80*nrow(df), replace = FALSE)
train <- df[smp_size,]
test <- df[-smp_size,]

```

# Step 2
Logistic regression model created and predicted on the yes/no test data plus the accuracy. 
```{r}
#a
fit.glm <- glm(default ~ income + balance, data = Default, family = "binomial")

#b
summary(fit.glm)
#c
pred <- predict(fit.glm, type = "response")
pred_value <- ifelse(pred > .5, 'Yes', 'No')

#d
acc1 <- mean(pred_value == Default$default)
print(paste("Accuracy: ", acc1))

```

# Step 3
Decision tree model created and summary was ran, and so was the accuracy that was computed. 
```{r}
library(rpart)
#a
tree_df <- rpart(default~., data= train, method = "class")
#b
summary(tree_df)
#c
pred_tree <- predict(tree_df, type ="class")

acc2 <- mean(pred_tree == train$default)
print(paste("Accuracy: ", acc2))
```

#Step 4
Here were display the tree that was created to show the decisions that were taken by the tree. 
```{r}
plot(tree_df, uniform = TRUE)
text(tree_df, use.n=TRUE, all=TRUE, cex=.6)
```


#### Problem 2 ####

#Step 1 

Setting up the data ones more, reading from the csv file saved on the desktop computer. loading the data into R and then attaching it. however it game a lot of errors with hidden data. Removing the X column, and setting train and test with 80%. 
```{r}
#a
Heart <- read.csv(file = '/Users/trishalvarma/Desktop/Heart.csv')
#b

#attach funciton giving error on masked objects. can remove the # for testing purposes. 

#attach(Heart) 

#c
df_heart <- Heart
df_heart <- subset(df_heart, select = -c(X))
# df_heart <- na.omit(df_heart)
na.rm = TRUE
#d
names(df_heart)
set.seed(2017)
smp_heart <- sample(1:nrow(df_heart), .80*nrow(df_heart), replace = FALSE)
heart.train <- df_heart[smp_heart,]
heart.test <- df_heart[-smp_heart,]
```

# Step 2 
Creating a logistical regression model on the training data where AHD is predicted by all other variable. Computed the accuracy.
```{r}
#a. 
heart.train$AHD <- as.factor(heart.train$AHD)
heart.test$AHD <- as.factor(heart.test$AHD)
heart_glm <- glm(AHD ~.-AHD, data = heart.test, family = "binomial")

#b
summary(heart_glm)

#c
heart.pred <- predict(heart_glm, type = "response")
heart_value <- ifelse(heart.pred > .5, 'Yes', 'No')

#d
heartacc <- mean(heart_value == heart.test$AHD)
print(paste("Accuracy: ", heartacc))
```

# Step 3
Creating a decision tree model on the training data set and running the summary. 
computed the accuracy. 
```{r}
#a
tree_df2 <- rpart(AHD~., data = heart.train, method = "class")
#b
summary(tree_df2)
#c
pred_heart.tree <- predict(tree_df2, type ="class")
#d
heart_treeacc <- mean(pred_heart.tree == heart.train$AHD)
print(paste("Accuracy: ", heart_treeacc))
```

# Step 4
Here we display the tree that were created. 
```{r}
#a
plot(tree_df2, uniform = TRUE)
#b
text(tree_df2, use.n = TRUE, all=TRUE, cex=.6)

```
# Step 5
Cross validation function used to get the best n value and plottedß on a 1x2 format.
```{r}
#a 
tree <- tree(heart.train$AHD~., data = heart.train)
cv_tree <- cv.tree(tree)
#b
cv_tree$dev
cv_tree$size
#c
par(mfrow = c(1,2))
plot(cv_tree$size, cv_tree$dev, type='b')
plot(cv_tree$k, cv_tree$dev, type = 'b')
```

# Step 6
Pruned the tree with the best n value we found from step 5. 
```{r}
#a. 
tree_pruned <- prune.tree(tree, best=5)
#b
plot(tree_pruned)
text(tree_pruned, pretty = 0)

```

# Step 7
predicting on the pruned tree, and computing the accuracy of it. 
```{r}
#a.

prune_pred <- predict(tree_pruned, newdata = heart.test)

#b.
prune_acc <- mean(prune_pred)
print(paste("Accuracy: ", prune_acc))
```

Questions: 

Problem 1 
1. Income and Balance were important, and default, student were not. 
2. Accuracy for the logistical regression was 97.37% and for the decision tree was 97.41% 
3. One reason can be that since it is a greedy algorithm, and once a decision has been made, it doesn't reconsider earlier steps. The other reason could be cause it maybe more pure than the other, so it can have two no. 
4. if balance > 1800
    then yes
    else no 
5. The accuracy is pretty high in both the tree and the regression, so there is no need to prune the tree. 

Problem 2
1. None of the variables as shows were important in the regression model. As their Pr(>|z|) value is higher than .5. 
2. The AHD data was used to create the decision tree. 
3. Logistic accuracy was 80.32%, while the Decision tree accuracy was 86.36% 
4. Pruned tree resulted in a value of 50% 
5. Judging by just the accuracy of the data calculated, the decision tree accuracy may be more important to a doctor than the others. 

